---
title: "worksheet"
author: "Ahmet Faruk Ulutaş"
output:
  html_document:
    theme: united
date: "2022-10-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First thing is to call the libraries need in the worksheet
##call ggplot2 and dplry

```{r message=FALSE}
library(ggplot2)
library(dplyr)
```

## msleep contains ....
## msleep contains the mammals sleep dataset

```{r message=FALSE}
data(msleep)
?msleep
head(msleep)
```

## how many rows and columns the data have?
## The data have 83 rows and 11 columns

```{r message=FALSE}
sleepData <- select(msleep, name, sleep_total)
head(sleepData)
```

## what does "select" do? what does sleepData contains?
## Columns' names and kinds are used to choose subsets (sleep total in this code). A table in sleepData has 83 rows and 2 columns (name of the animals and their total amount of sleep in hours)

```{r message=FALSE}
head(select(msleep, starts_with("sl")))
```

## does the above statement works on rows or columns? which dataset it uses? Explain.
## The code above functions on columns. It utilizes the msleep dataset but only displays the "sl"-named columns (sleep total, sleep rem, and sleep cycle).

## Another important function in dplyr is to filter... 
## Another important function in dplyr is to filter subsets rows using column values.

```{r message=FALSE}
filter(msleep, sleep_total >= 16) 
filter(msleep, sleep_total >= 16, bodywt >= 1) 
filter(msleep, order %in% c("Perissodactyla", "Primates"))
```

## what does %in% do? what happens if you do not provide a vector next to it?
## If a match exists or not, %in% returns a logical vector for the left operand. If a vector is not given next to it, a syntax error occurs.

## pipe operator

```{r message=FALSE}
msleep %>%
  select(name, sleep_total) %>%
  head()
```

## what does %>% do? Compare with select(msleep, name, sleep_total)
## The expression %>% forwards an object to a function or call expression. head(select(msleep, name, sleep_total)) is the same as the code above.

## when you use %>% what is it doing?
## Msleep is piped into the function select, which is subsequently piped into the function head via %>%.

## arrange()

```{r message=FALSE}
msleep %>% 
  select(name, order, sleep_total) %>% 
  arrange(order, sleep_total) %>% 
  head()
```
## what does arrange do?
## Arrange groups rows according to column values. The rows are arranged here according to order and sleep_total.

## mutate()

```{r message=FALSE}
msleep %>% 
  mutate(rem_proportion = sleep_rem / sleep_total) %>% 
  head()
```

## what does mutate do?
## Create, modify, and remove are all possible using mutate. Here, it adds a "rem proportion" field to the msleep table.

## summarise()

```{r message=FALSE}
msleep %>% 
  summarise(avg_sleep = mean(sleep_total), 
            min_sleep = min(sleep_total),
            max_sleep = max(sleep_total),
            total = n())
```

## explain what does summarize do?
## A new data frame is created by summarise. Here, it generates a data frame with the animals' average, minimal, and maximal sleep durations.

## Write a short paragraph explaining the results from msleep, e.g., how much the animals sleep, what is the minimum and max. amount of sleep etc.
## The average amount of sleep for animals is 10.4 hours. The giraffe requires 1.9 hours of sleep every night as a minimum. The small brown bat sleeps for a maximum of 19.9 hours every night.
